---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(MOFA2)
source(here::here("code/funcs.R"))

set.seed(2022)
conflicted::conflicts_prefer(dplyr::select(), dplyr::rename(), dplyr::filter())
```

Testing the example from tutorial

```{r}
set.seed(2020)

# set number of samples and time points
N <- 200
time <- seq(0,1,length.out = N)

# generate example data
dd <- make_example_data(sample_cov = time, n_samples = N,
                        n_factors = 4, n_features = 200, n_views = 4,
                        lscales = c(0.5, 0.2, 0, 0))
# input data
data <- dd$data

# covariate matrix with samples in columns
time <- dd$sample_cov
rownames(time) <- "time"
```

Looks like I'll need two "views" (matrices) in a special datatype, one for community abundance, one for metabolomics, maybe a third for gene content? \## Read in data

```{r}
metadata <- read_metaG_metadata()
```

### Community composition

```{r}
percent_abund <- read_rds("results/percent_abund.rds")

rel.abund <- percent_abund %>% 
  dplyr::select(taxonomy = "classification", Domain, Phylum, Class, Order, Family, Genus, Species, taxonomy_id = "genome", percent_abund, sample) %>% 
  pivot_wider(values_from = "percent_abund", names_from = "sample",values_fill = 0) %>% 
  mutate(taxonomy = str_remove_all(taxonomy,"(;[a-z]__)+$"),
         taxonomy = if_else(taxonomy == "Unclassified Bacteria", "d__Bacteria", taxonomy),
         taxonomy = if_else(taxonomy == "Unclassified" | is.na(taxonomy) , "", taxonomy),
         taxonomy = paste0("r__root;",taxonomy),
         taxonomy = str_remove(taxonomy, ";$"))
  
rel_abund_long <- rel.abund %>% 
  pivot_longer(any_of(metadata$sample), names_to = "sample",values_to = "rel_abund")

vegan_abund <- percent_abund %>% 
  ungroup() %>% 
  dplyr::select(genome,sample, percent_abund) %>% 
  distinct() %>% 
  pivot_wider(names_from = sample, values_from = percent_abund, values_fill = 0) %>% 
  column_to_rownames("genome") %>% 
  t()
```

### Metabolomics

```{r}
metabolite_abund <- read_rds("data/metabolomics/combined.rds")

metabolite_info <- read_rds("data/metabolomics/metabolite_info_combined.rds")

named_metabolites <- metabolite_info %>%
  select(`Compound Name`, `Feature Name or Formula`) %>%
  filter(!is.na(`Compound Name`) | !str_detect(`Feature Name or Formula`, "^UNK_")) %>% 
  mutate(`Compound Name` = if_else(is.na(`Compound Name`), `Feature Name or Formula`, `Compound Name`))

named_metabolite_abund <- metabolite_abund %>%
  rownames_to_column("compound") %>%
  filter(compound %in% named_metabolites$`Feature Name or Formula`) %>%
  column_to_rownames("compound")

metabolite_abund_vegan <- read_rds("data/metabolomics/combined_vegan.rds")
```

### Gene abundance

```{r}
# Read in KO relative abundance
ko_abund <- read_rds("results/bin_ko_abund.rds")

ko_defs <- read_rds("results/ko_defs.rds")

#read_rds("results/bin_ko_abund_per_bin.rds")
```

## Prep & merge data

```{r}
mofa_tax <- rel_abund_long %>% 
  filter(!is.na(Class),
         !is.na(Species)) %>% 
  mutate(taxonomy = str_glue("{Domain}; {Phylum}; {Class}; {Order}; {Family}; {Genus}; {Species}")) %>% 
  group_by(sample, Domain, Phylum, Class, Order, Family, Genus, Species, taxonomy) %>% 
  summarise(rel_abund = sum(rel_abund)) %>% 
  ungroup() %>% 
  select(sample, feature = "taxonomy", value = "rel_abund") %>% 
  filter(value > 0) %>% 
  mutate(value = log(value),
         view = "tax_abund",
         group = NA)

# Look at distribution of tax abund data
mofa_tax %>% ggplot(aes(value)) +
  geom_histogram() #+
  #scale_x_log10()

mofa_metabolites <- named_metabolite_abund %>%
  rownames_to_column("feature") %>%
  pivot_longer(-feature,names_to = "sample", values_to = "value") %>%
  filter(value > 0) %>%
  mutate(value = log(value),
  group = NA,
  view = "metabolites")

# Look at distribution of metabolite abund data
mofa_metabolites %>% ggplot(aes(value)) +
  geom_histogram() #+
  #scale_x_log10()
  
# KEGG KO abund

ko_names <- 

mofa_ko <- ko_abund %>% 
  select(sample, feature = "ko", value = "ko_abund") %>% 
  mutate(group = NA,
         view = "ko_abund",
         value = log(value)) %>% 
  filter(value > 0)
  
mofa_combined <- bind_rows(mofa_tax, 
                           mofa_metabolites, 
                           mofa_ko
                           ) %>% 
  select(-group) #%>%
  #left_join(metadata %>% select(sample, group = "replicate"))

```

## MOFA analysis

```{r}
mofa_obj <- create_mofa_from_df(mofa_combined)

plot_data_overview(mofa_obj)
```

```{r}
model_opts <- get_default_model_options(mofa_obj)
model_opts$num_factors <- 10

mofa <- prepare_mofa(mofa_obj, model_options = model_opts)

mofa_trained <- run_mofa(mofa,use_basilisk = TRUE)

samples_metadata(mofa_trained) <- metadata

fs::dir_create("data/mofa")
write_rds(mofa_trained,"data/mofa/mofa_trained.rds")
```

```{r}
plot_variance_explained(mofa_trained, plot_total = T)[[2]]
```

```{r}
plot_variance_explained(mofa_trained, max_r2=15)
```

```{r}
MOFA2::plot_factors(mofa_trained, 
  factors = c(1,5), 
   color_by = "plot_categories",
  # shape_by = "gut_section",
  #color_by = "gut_section",
  #color_by = "days_after_transplant",
  #shape_by = "plot_categories",
  dot_size = 4
) #+ scale_fill_manual(values=category.colors)
```

```{r}



plot_factor(mofa_trained, 
  factor = 1, 
  color_by = "gut_section", 
  dot_size = 4,
  dodge = TRUE,
  stroke = 0.4,
  add_violin = T,
  add_boxplot = T
) +
  #scale_fill_manual(values=category.colors) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )




plot_factor(mofa_trained, 
  factor = 5, 
  color_by = "plot_categories", 
  dot_size = 4,
  dodge = TRUE,
  stroke = 0.4,
  add_violin = T,
  add_boxplot = T
) +
  #scale_fill_manual(values=category.colors) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

```

```{r}
plot_weights_fn <- function(mofa, factor=1, view=1, nfeatures=10) {
  p1 <- plot_weights(mofa, 
    factors = factor, 
    view = view,
    nfeatures = nfeatures,
    text_size = 4
  )
  
  p2 <- plot_top_weights(mofa, 
    factors = factor, 
    view = view,
    nfeatures = nfeatures
  )
  
  p <- cowplot::plot_grid(plotlist=list(p1,p2), nrow=1)
  return(p)
}
```

```{r}
plot_weights_fn(mofa_trained, factor=1, view="tax_abund", nfeatures=8)
plot_weights_fn(mofa_trained, factor=1, view="metabolites", nfeatures=8)

factor1_weights <- get_weights(mofa_trained,
                               factors = ,
                               as.data.frame = TRUE) %>% 
  left_join(metabolite_info %>% 
              select(feature = "Feature Name or Formula", 
                     compound_name = "Compound Name",
                     metab_super_class = "super_class",
                     metab_main_class = "main_class",
                     metab_sub_class = "sub_class") %>% 
              mutate(view = "metabolites") %>% 
              distinct()
            ) %>% 
  left_join(rel_abund_long %>% 
              select(feature = "taxonomy_id", 
                    Domain, Phylum, Class, Order, Family, Genus, Species, taxonomy) %>% 
              mutate(view = "tax_abund") %>% 
              distinct()
            ) %>% 
  left_join(ko_defs %>% 
              select(feature = "ko", 
                    ko_definition) %>% 
              mutate(view = "ko_abund") %>% 
              distinct()
            )

factor5_weights <- get_weights(mofa_trained,
                               factors = 5,
                               as.data.frame = TRUE) %>% 
  left_join(metabolite_info %>% 
              select(feature = "Feature Name or Formula", 
                     compound_name = "Compound Name",
                     metab_super_class = "super_class",
                     metab_main_class = "main_class",
                     metab_sub_class = "sub_class") %>% 
              mutate(view = "metabolites") %>% 
              distinct()
            ) %>% 
  left_join(rel_abund_long %>% 
              select(feature = "taxonomy_id", 
                    Domain, Phylum, Class, Order, Family, Genus, Species, taxonomy) %>% 
              mutate(view = "tax_abund") %>% 
              distinct()
            ) %>% 
  left_join(ko_defs %>% 
              select(feature = "ko", 
                    ko_definition) %>% 
              mutate(view = "ko_abund") %>% 
              distinct()
            )



plot_weights_fn(mofa_trained, factor=5, view="tax_abund", nfeatures=8)
plot_weights_fn(mofa_trained, factor=5, view="metabolites", nfeatures=20)
```

```{r}
plot_data_heatmap(mofa_trained, 
  factor = 5, 
  view = "tax_abund", 
  features = 20,
  denoise = FALSE,
  cluster_rows = T, cluster_cols = F,
  show_colnames = F, show_rownames = T,
  annotation_samples = "plot_categories",  
  #annotation_colors = list("Category"=category.colors), 
  annotation_legend = F,
  scale = "row"
)
```

```{r}
plot_data_scatter(mofa_trained, 
  factor = 5, 
  view = "tax_abund", 
  features = 4,
  dot_size = 3,
  color_by = "plot_categories",
  legend = F
)


```

```{r}


```

## MEFEISTO

```{r}
mofa_obj <- create_mofa_from_df(mofa_combined)

samples_metadata(mofa_obj) <- metadata %>% 
  #dplyr::rename(group = "replicate") %>% 
  mutate(numeric_gut_section = as.numeric(gut_section),
         numeric_days_since_transplant = as.numeric(days_after_transplant),
         numeric_days_since_transplant = if_else(plot_categories == "Naive", 0, numeric_days_since_transplant))

mofa_obj_spatial <- set_covariates(mofa_obj,
                                   covariates =  c("numeric_gut_section","numeric_days_since_transplant"))

plot_data_overview(mofa_obj_spatial)
```

```{r}

data_opts <- get_default_data_options(mofa_obj_spatial)


model_opts_spatial <- get_default_model_options(mofa_obj_spatial)
model_opts_spatial$num_factors <- 15

mefisto_opts <- get_default_mefisto_options(mofa_obj_spatial)

mofa_spatial <- prepare_mofa(mofa_obj_spatial, 
                             model_options = model_opts_spatial,
                             data_options = data_opts,
                             mefisto_options = mefisto_opts,)

mofa_spatial_trained <- run_mofa(mofa_spatial,use_basilisk = TRUE)



fs::dir_create("data/mofa")
write_rds(mofa_spatial_trained,"data/mofa/mofa_spatial_trained.rds")
```

```{r}
plot_variance_explained(mofa_spatial_trained, plot_total = T)[[2]]
```

```{r}
plot_variance_explained(mofa_spatial_trained, max_r2=15)
```

```{r}
r <- plot_factor_cor(mofa_spatial_trained)
```

From [this](https://raw.githack.com/bioFAM/MEFISTO_tutorials/master/MEFISTO_spatial.html) tutorial:

> The MOFA model has learnt scale parameters for each factor, which give us an indication of the smoothness per factor along the covariate (here space) and are between 0 and 1. A scale of 0 means that the factor captures variation independent of space, a value close to 1 tells us that this factor varys very smoothly along space.

```{r}
get_scales(mofa_spatial_trained)
```

```{r}
plot_factors_vs_cov(mofa_spatial_trained) 
```



```{r}

factor_pair <- c(1,5)


MOFA2::plot_factors(mofa_spatial_trained, 
  factors = factor_pair, 
  color_by = "plot_categories",
  # shape_by = "gut_section",
  #color_by = "gut_section",
  #color_by = "days_after_transplant",
  #shape_by = "plot_categories",
  dot_size = 4
) #+ scale_fill_manual(values=category.colors)


MOFA2::plot_factors(mofa_spatial_trained, 
  factors = factor_pair, 
  #color_by = "plot_categories",
  # shape_by = "gut_section",
  color_by = "gut_section",
  #color_by = "days_after_transplant",
  #shape_by = "plot_categories",
  dot_size = 4
) #+ scale_fill_manual(values=category.colors)
```

```{r}

factors <- 1:10


plot_factor_box <- function(fact_num, color_param){
  plt <- (plot_factor(mofa_spatial_trained, 
  factor = fact_num, 
  color_by = color_param, 
  dot_size = 4,
  dodge = TRUE,
  stroke = 0.4,
  add_violin = T,
  add_boxplot = T
) +
  #scale_fill_manual(values=category.colors) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  ))
  
  print(plt)
}

walk2(factors, "plot_categories", .f = plot_factor_box)
walk2(factors, "gut_section", .f = plot_factor_box)

```

```{r}
plot_weights_fn <- function(mofa, factor=1, view=1, nfeatures=10) {
  p1 <- plot_weights(mofa, 
    factors = factor, 
    view = view,
    nfeatures = nfeatures,
    text_size = 4
  )
  
  p2 <- plot_top_weights(mofa, 
    factors = factor, 
    view = view,
    nfeatures = nfeatures
  )
  
  p <- cowplot::plot_grid(plotlist=list(p1,p2), nrow=1)
  return(p)
}
```


```{r}
factor1_weights_spatial <- get_weights(mofa_spatial_trained,
                               factors = 1,
                               as.data.frame = TRUE) %>% 
  left_join(metabolite_info %>% 
              select(feature = "Feature Name or Formula", 
                     compound_name = "Compound Name",
                     metab_super_class = "super_class",
                     metab_main_class = "main_class",
                     metab_sub_class = "sub_class") %>% 
              mutate(view = "metabolites") %>% 
              distinct()
            ) %>% 
  # left_join(rel_abund_long %>% 
  #             select(feature = "taxonomy_id", 
  #                   Domain, Phylum, Class, Order, Family, Genus, Species, taxonomy) %>% 
  #             mutate(view = "tax_abund") %>% 
  #             distinct()
  #           ) %>% 
  left_join(ko_defs %>% 
              select(feature = "ko", 
                    ko_definition) %>% 
              mutate(view = "ko_abund") %>% 
              distinct()
            )


factor8_weights_spatial <- get_weights(mofa_spatial_trained,
                               factors = 8,
                               as.data.frame = TRUE) %>% 
  left_join(metabolite_info %>% 
              select(feature = "Feature Name or Formula", 
                     compound_name = "Compound Name",
                     metab_super_class = "super_class",
                     metab_main_class = "main_class",
                     metab_sub_class = "sub_class") %>% 
              mutate(view = "metabolites") %>% 
              distinct()
            ) %>% 
  left_join(rel_abund_long %>% 
              select(feature = "taxonomy_id", 
                    Domain, Phylum, Class, Order, Family, Genus, Species, taxonomy) %>% 
              mutate(view = "tax_abund") %>% 
              distinct()
            ) %>% 
  left_join(ko_defs %>% 
              select(feature = "ko", 
                    ko_definition) %>% 
              mutate(view = "ko_abund") %>% 
              distinct()
            )
```


```{r}
plot_weights_fn(mofa_spatial_trained, factor=1, view="tax_abund", nfeatures=8)
plot_weights_fn(mofa_spatial_trained, factor=1, view="metabolites", nfeatures=8)


plot_weights_fn(mofa_spatial_trained, factor=5, view="tax_abund", nfeatures=8)
plot_weights_fn(mofa_spatial_trained, factor=5, view="metabolites", nfeatures=20)
```

```{r}
plot_data_heatmap(mofa_trained, 
  factor = 5, 
  view = "tax_abund", 
  features = 20,
  denoise = FALSE,
  cluster_rows = T, cluster_cols = F,
  show_colnames = F, show_rownames = T,
  annotation_samples = "plot_categories",  
  #annotation_colors = list("Category"=category.colors), 
  annotation_legend = F,
  scale = "row"
)
```

```{r}
plot_data_scatter(mofa_trained, 
  factor = 5, 
  view = "tax_abund", 
  features = 4,
  dot_size = 3,
  color_by = "plot_categories",
  legend = F
)


```
