---
title: "upset plot of GVHD metabolites differentially abundant between syn & allo"
output: html_notebook
---

This 'setup' chunk gets automatically run the first time code is executed from the document
```{r setup}
knitr::opts_knit$set(root.dir = here::here("~/GVHD")) # this sets the working directory, would need to be changed if paths are different

# Load the packages used
library(tidyverse)
library(UpSetR)
```


## Locations syn/allo

Read in the metabolite table
```{r}
dysregulated_mets <- readxl::read_xlsx("data/metabolomics/dysregulated_mets_bylocation.xlsx")
```

Convert to a presence absence table (0/1) with groups/categories/labels as columns, and the features/genes/metabolites/etc as rows
```{r}
met_pres_abs <- dysregulated_mets %>% 
  pivot_longer(everything(), names_to = "gut_section",
               values_to = "metabolite") %>% 
  filter(!is.na(metabolite)) %>% 
  mutate(met_pres = 1) %>% 
  pivot_wider(values_from = met_pres, names_from = gut_section,values_fill = 0) %>% 
  write_tsv("data/metabolomics/dysregulated_mets_bylocation_pres_abs.tsv") %>% 
  column_to_rownames("metabolite")
```

Make the actual plot
```{r}
# this and the dev.off() lines are used for saving to a pdf
pdf(file="~/GVHD/results/dysregulated_metabolites_upset.pdf",width = 4, height = 3) # change pdf path here

upset(met_pres_abs,order.by = "freq") # changed the order to be ordered by the intersection size

dev.off()
```

Make as a heatmap of overlaps instead
```{r}

met_pres_abs_long <- met_pres_abs %>% 
  rownames_to_column("met") %>% 
  pivot_longer(-met, names_to = "gut_section", values_to = "pres")

combinations <- utils::combn(colnames(met_pres_abs),2,simplify = TRUE) %>% t() %>% as.data.frame() %>% 
  dplyr::rename(sec1 = "V1", sec2 = "V2") %>% 
  left_join(met_pres_abs_long %>% dplyr::rename(sec1 = "gut_section", pres1 = "pres")) %>% 
  left_join(met_pres_abs_long %>% dplyr::rename(sec2 = "gut_section", pres2 = "pres"))


overlaps <- combinations %>% 
  filter(pres1 == 1 & pres2 == 1) %>% 
  group_by(sec1, sec2) %>% 
  summarise(overlaps = n())

overlaps %>% 
  ggplot(aes(sec1, sec2, label = overlaps, fill = overlaps)) +
  geom_tile() + 
  geom_text(size = 6) +
  theme_bw() +
  scale_fill_viridis_c()
```


## By days & locations diff abund syn/allo

Read in the metabolite table
```{r}
dysregulated_mets <- read_tsv("data/metabolomics/diff_abund_metabolites_by_day.txt") %>% 
  column_to_rownames("all")
```
Make the actual plot
```{r}
# this and the dev.off() lines are used for saving to a pdf
pdf(file="~/GVHD/results/dysregulated_metabolites_upset_by_days.pdf",width = 8, height = 5) # change pdf path here

upset(dysregulated_mets,order.by = "freq",nsets = 8) # changed the order to be ordered by the intersection size

dev.off()
```

Make as a heatmap of overlaps instead
```{r}

met_pres_abs_long <- dysregulated_mets %>% 
  rownames_to_column("met") %>% 
  pivot_longer(-met, names_to = "gut_section", values_to = "pres")

combinations <- utils::combn(colnames(dysregulated_mets),2,simplify = TRUE) %>% t() %>% as.data.frame() %>% 
  dplyr::rename(sec1 = "V1", sec2 = "V2") %>% 
  left_join(met_pres_abs_long %>% dplyr::rename(sec1 = "gut_section", pres1 = "pres")) %>% 
  left_join(met_pres_abs_long %>% dplyr::rename(sec2 = "gut_section", pres2 = "pres"))


dist <- dist(dysregulated_mets %>% t()) 

overlaps <- combinations %>% 
  filter(pres1 == 1 & pres2 == 1) %>% 
  group_by(sec1, sec2) %>% 
  summarise(overlaps = n())

diff_abund_mets_hclust <- hclust(dist)

 plot(diff_abund_mets_hclust)

order <- diff_abund_mets_hclust$labels[diff_abund_mets_hclust$order]

overlaps %>% 
  
  ggplot(aes(reorder(sec1, ), relevel(as.factor(sec2),diff_abund_mets_hclust$labels[diff_abund_mets_hclust$order]), label = overlaps, fill = overlaps)) +
  geom_tile() + 
  geom_text(size = 6) +
  theme_bw() +
  scale_fill_viridis_c()
```




